### 2020-08-26 thesis/appendix_1.tex
replace montreal adn quebec by accented versions

\begin{table}[h!]
\centering
\caption{Description and data sources for all variables used in the Random Forest - Cellular Automata model.}
\label{tab:variables}
\begin{tabular}{lll}
\hline
\textbf{Variable} & \textbf{Format} & \textbf{Source}  \\ 
\hline
Distance from urban land & \multirow{3}{*}{Raster} & \multirow{2}{*}{Generated from land cover data} \\ \cline{1-1}
Size of forest patch &  &  \\ \cline{1-1} \cline{3-3} 
Elevation &  & \begin{tabular}[c]{@{}l@{}}SRTM 30m from \\ Google Earth Engine \\ data library\end{tabular} \\ \hline
Population change & \multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}Tabular data joined to vector \\ data and rasterized\end{tabular}} & \multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}Canadian Census for \\ 1991, 2001 and 2011\end{tabular}} \\ \cline{1-1}
Income &  &  \\ 
\hline
\end{tabular}
\end{table}

>>>>

\begin{table}[h!]
\centering
\caption{Description and data sources for all variables used in the Random Forest - Cellular Automata model.}
\label{tab:variables}
\begin{tabular}{lll}
\hline
\textbf{Variable} & \textbf{Format} & \textbf{Source}  \\ 
\hline
Distance from urban land & \multirow{3}{*}{Raster} & \multirow{2}{*}{Generated from land cover data} \\ \cline{1-1}
Size of forest patch &  &  \\ \cline{1-1} \cline{3-3} 
Elevation &  & \begin{tabular}[c]{@{}l@{}}SRTM 30m from \\ Google Earth Engine \\ data library\end{tabular} \\ \hline
Population change & \multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}Tabular data joined to vector \\ data and rasterized\end{tabular}} & \multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}Canadian Census for \\ 1991, 2001 and 2011\end{tabular}} \\ \cline{1-1}
Income &  &  \\ \hline
Road network & Vector data (rasterized) & Adresses Québec \\ 
\hline
\end{tabular}
\end{table}

### 2020-08-26 thesis/chap_1.tex
no repro appendix - stick with readme
remove todo
replace montreal adn quebec by accented versions

A note should be added on data availability and reproducibility. A few datasets were not obtained from open sources. These datasets are marked in red in figure \ref{fig:workflow1}. The Canadian census data was obtained through the University of Toronto's CHASS (Computing in the Humanities and Social Sciences) Data Centre, which requires institutional access. This makes the Random Forest analysis not reproducible. In addition, data on protected areas in Montérégie was obtained  via the RMN (Réseaux des milieux naturels protégés du Québec) under a limited license, which makes the Land use change modeling via ST-Sim not fully reproducible. However, the results of those steps are made open, and the subsequent steps (habitat modeling, connectivity analysis) are reproducible via the use of various software ( i.e. docker, renv, and more). Please consult the appendix on reproducibility for the steps to take to reproduce those results. \\ %TODO write appendix on reproducibility

@@

This data is not without limitations: compared to other data products - such as the Quebec Ministry of the Environment land use dataset or the Statistique Quebec land accounting dataset - the AAFC dataset is not very detailed: only a few land use categories are identified, and important information for suitability analysis such as forest age and forest density are missing.

Another limitation is the coordinate reference system (CRS) in which this dataset is available: it is projected in a UTM projection system, which is not the standard CRS used by other data products that usually covers Quebec and emanates from open data portals in the province. This makes direct comparison (pixel by pixel) to those others datasets difficult. Finally, the AAFC data does not differentiate between road types, which has implications for habitat suitability analysis (see the habitat suitability analysis section).\\

@@

%---------------------------------------------------------------------------------------------------------------------------------------------------
% methods figures %TODO better captions

>>>>

A note should be added on data availability and reproducibility. A few datasets were not obtained from open sources. These datasets are marked in red in figure \ref{fig:workflow1}. The Canadian census data was obtained through the University of Toronto's CHASS (Computing in the Humanities and Social Sciences) Data Centre, which requires institutional access. This makes the Random Forest analysis not reproducible. In addition, data on protected areas in Montérégie was obtained  via the RMN (Réseaux des milieux naturels protégés du Québec) under a limited license, which makes the Land use change modeling via ST-Sim not fully reproducible. However, the results of those steps are made open, and the subsequent steps (habitat modeling, connectivity analysis) are reproducible via the use of various software ( i.e. docker, renv, and more). \\

@@

This data is not without limitations: compared to other data products - such as the Québec Ministry of the Environment land use dataset or the Statistique Québec land accounting dataset - the AAFC dataset is not very detailed: only a few land use categories are identified, and important information for suitability analysis such as forest age and forest density are missing.

Another limitation is the coordinate reference system (CRS) in which this dataset is available: it is projected in a UTM projection system, which is not the standard CRS used by other data products that usually covers Québec and emanates from open data portals in the province. This makes direct comparison (pixel by pixel) to those others datasets difficult. Finally, the AAFC data does not differentiate between road types, which has implications for habitat suitability analysis (see the habitat suitability analysis section).\\

@@

### 2020-08-26 thesis/chap_2.tex
replace montreal adn quebec by accented versions

Here, we do not present a scenario design exercise, but a community workshop with a consensus building and participatory mapping exercise, whose results were translated into simple scenarios in an integrated land use, climate change, and connectivity model (developed in chapter 1). We improve our understanding of the obstacles and opportunities to connectivity conservation in the region, and provide a good starting ground for an iterative process of scenario building. We focused on the region of Montérégie, in southern Quebec, where connectivity conservation has become the focus of land use planning decisions. This project has been developed in collaboration with the non-profit NAQ (Nature action Quebec) in the context of a multi-stakeholder conservation project, the PADF (Plan d’Aménagement Durable des Forêts). The PADF aims at building consensus on priorities for the management and possible conservation of forested land in the region.

@@

The presence of species at risk was identified as an important factor, along with other elements such as social factors (available economic means and the level of local interest in conservation) and characteristics of the link itself, including the functionality of the corridor. Consideration of threats such as the level of fragmentation was also noted. Finally, the fact that the environment is to be protected or restored was also deemed important. Interestingly, regulations are among the factors that were cited less often, such as Quebec's PRMHH ("Plans régionaux des milieux humides et hydriques", or regional wetlands conservation plans), and migration routes. \\

@@

The simple approach to the sparing vs sharing question that is brought about by our scenario overlooks the question of scale. At the scale of Monteregie, it is unclear to which extent the reduction in agricultural land can be promoted by the regional government. Like all modern farming systems, Quebec's agriculture is embedded in a global market which might make it likely to prioritize global trends over local interests. Therefore, we are overlooking important variables for the design of credible alternative scenarios for conservation in the region, variables that are more likely to be elucidated in a participatory setting like the one we used.

>>>>

Here, we do not present a scenario design exercise, but a community workshop with a consensus building and participatory mapping exercise, whose results were translated into simple scenarios in an integrated land use, climate change, and connectivity model (developed in chapter 1). We improve our understanding of the obstacles and opportunities to connectivity conservation in the region, and provide a good starting ground for an iterative process of scenario building. We focused on the region of Montérégie, in southern Québec, where connectivity conservation has become the focus of land use planning decisions. This project has been developed in collaboration with the non-profit NAQ (Nature action Québec) in the context of a multi-stakeholder conservation project, the PADF (Plan d’Aménagement Durable des Forêts). The PADF aims at building consensus on priorities for the management and possible conservation of forested land in the region.

@@

The presence of species at risk was identified as an important factor, along with other elements such as social factors (available economic means and the level of local interest in conservation) and characteristics of the link itself, including the functionality of the corridor. Consideration of threats such as the level of fragmentation was also noted. Finally, the fact that the environment is to be protected or restored was also deemed important. Interestingly, regulations are among the factors that were cited less often, such as Québec's PRMHH ("Plans régionaux des milieux humides et hydriques", or regional wetlands conservation plans), and migration routes. \\

@@

The simple approach to the sparing vs sharing question that is brought about by our scenario overlooks the question of scale. At the scale of Monteregie, it is unclear to which extent the reduction in agricultural land can be promoted by the regional government. Like all modern farming systems, Québec's agriculture is embedded in a global market which might make it likely to prioritize global trends over local interests. Therefore, we are overlooking important variables for the design of credible alternative scenarios for conservation in the region, variables that are more likely to be elucidated in a participatory setting like the one we used.

### 2020-08-26 thesis/thesis.tex
replace montreal adn quebec by accented versions

\SetAuthor{\large{\textit{Valentin Lucet}}}%
\SetDegreeType{Master of Science\vspace{0cm}}%
\SetDepartment{Department of Biology\vspace{-2.1cm}}%
\SetUniversity{McGill University}%
\SetUniversityAddr{Montreal, Quebec}%
\SetThesisDate{August 2020}%
\SetRequirements{A thesis submitted to McGill University in partial fulfillment \\of the requirements of the degree of Master of Science}%

@@

Ecological connectivity is the extent to which the landscape supports the movements of organisms \citep{gonzalez_spatial_2017}, and is paramount for the resilience of both populations and ecosystem services \citep{mitchell_monteregie_2015} in heterogeneous and fragmented landscapes. Land use changes such as urban sprawl can cause deforestation, fragmenting habitats, and slowly eroding ecological connectivity. Many urban landscapes are experiencing uncontrolled urban sprawl and have suffered losses in connectivity and ecosystem services in consequence. Examples include cities like Barcelona \citep{marulli_gis_2005}, New York City \citep{mcphearson_urban_2014}, and also Montreal \citep{dupras_urban_2015}.  The drivers behind those land use changes are complex and understanding them is an obstacle to conservation planning  \citep{worboys_connectivity_2010}, including connectivity conservation planning. \\
Connectivity conservation planning (CCP), refers to the enterprise that engages multiple actors such as academics, NGOs, governmental bodies at different scales and in a common goal to conserve the ecological connectivity of the landscape. CCP methods usually entails modelling connectivity of the landscape of interest and using a prioritization method to determine conservation priorities, and modify land use practices to conserve or improve connectivity. This study constitutes a contribution to an ongoing effort of CCP in Montérégie, in southern Quebec. In the following, we introduce the study region and describe previous efforts of CCP. We then point out ways in which CCP methods can be improved and how this thesis contributes to doing so. \\ \vspace{2em}

@@

Montérégie is one of the 17 administrative regions of the canadian province of Quebec. It is situated southeast of the city of Montreal, Québec, on the south bank of the Saint Lawrence, and is therefore located in the ecoregion of the Saint Lawrence Lowlands. Its superficy is of \SI{11852}{\km\squared} and is home to about $1.5$ million inhabitant. Agriculture is the predominant land use in the region, and about half of this superficy is covered by agricultural land \citep{isq_quebec_2020}. Statistique Quebec describes that between 1995 and 2008, 8\% of natural habitat and 4\% of agricultural lands has been artificialised. These changes are spread over the 15 MRC (“Municipalite Regionales de Comte”) making up Montérégie. MRCs are essentially sub-administrative regions at the level of which land use orientations are decided through two important documents: the PDZA (Plan de Développement de la Zone Agricole) and the SAD (Schéma d'Aménagement et de Développement). These MRCs are made up of 177 municipalities. In addition to MRCs, two First Nation reserves are present in Montérégie. \\

The eastern part of Montérégie is also under the jurisdiction of the Greater Montreal Area (GMA). The ecological connectivity of the GMA and its benefits as a provider of ecosystem services has recently been assessed in a report to the Quebec ministry of the environment (Rayfield 2018, unpublished). This study focused on identifying regions of highest connectivity, and therefore of highest priority for the conservation of biodiversity and ecosystem services. Other work by Rayfield et al. (2019, unpublished) has extended the analysis of connectivity to the whole of the Saint Lawrence Lowlands. Beyond connectivity the project of Monteregie Connection, spearheaded by an interdisciplinary team of researchers from McGill University, studied the provision of ecosystem services in the region and produced important knowledge for conservation planning. \\

>>>>

\SetAuthor{\large{\textit{Valentin Lucet}}}%
\SetDegreeType{Master of Science\vspace{0cm}}%
\SetDepartment{Department of Biology\vspace{-2.1cm}}%
\SetUniversity{McGill University}%
\SetUniversityAddr{Montréal, Québec}%
\SetThesisDate{August 2020}%
\SetRequirements{A thesis submitted to McGill University in partial fulfillment \\of the requirements of the degree of Master of Science}%

@@

Ecological connectivity is the extent to which the landscape supports the movements of organisms \citep{gonzalez_spatial_2017}, and is paramount for the resilience of both populations and ecosystem services \citep{mitchell_monteregie_2015} in heterogeneous and fragmented landscapes. Land use changes such as urban sprawl can cause deforestation, fragmenting habitats, and slowly eroding ecological connectivity. Many urban landscapes are experiencing uncontrolled urban sprawl and have suffered losses in connectivity and ecosystem services in consequence. Examples include cities like Barcelona \citep{marulli_gis_2005}, New York City \citep{mcphearson_urban_2014}, and also Montréal \citep{dupras_urban_2015}.  The drivers behind those land use changes are complex and understanding them is an obstacle to conservation planning  \citep{worboys_connectivity_2010}, including connectivity conservation planning. \\
Connectivity conservation planning (CCP), refers to the enterprise that engages multiple actors such as academics, NGOs, governmental bodies at different scales and in a common goal to conserve the ecological connectivity of the landscape. CCP methods usually entails modelling connectivity of the landscape of interest and using a prioritization method to determine conservation priorities, and modify land use practices to conserve or improve connectivity. This study constitutes a contribution to an ongoing effort of CCP in Montérégie, in southern Québec. In the following, we introduce the study region and describe previous efforts of CCP. We then point out ways in which CCP methods can be improved and how this thesis contributes to doing so. \\ \vspace{2em}

@@

Montérégie is one of the 17 administrative regions of the canadian province of Quebec. It is situated southeast of the city of Montréal, Québec, on the south bank of the Saint Lawrence, and is therefore located in the ecoregion of the Saint Lawrence Lowlands. Its superficy is of \SI{11852}{\km\squared} and is home to about $1.5$ million inhabitant. Agriculture is the predominant land use in the region, and about half of this superficy is covered by agricultural land \citep{isq_quebec_2020}. Statistique Quebec describes that between 1995 and 2008, 8\% of natural habitat and 4\% of agricultural lands has been artificialised. These changes are spread over the 15 MRC (“Municipalite Regionales de Comte”) making up Montérégie. MRCs are essentially sub-administrative regions at the level of which land use orientations are decided through two important documents: the PDZA (Plan de Développement de la Zone Agricole) and the SAD (Schéma d'Aménagement et de Développement). These MRCs are made up of 177 municipalities. In addition to MRCs, two First Nation reserves are present in Montérégie. \\

The eastern part of Montérégie is also under the jurisdiction of the Greater Montréal Area (GMA). The ecological connectivity of the GMA and its benefits as a provider of ecosystem services has recently been assessed in a report to the Québec ministry of the environment (Rayfield 2018, unpublished). This study focused on identifying regions of highest connectivity, and therefore of highest priority for the conservation of biodiversity and ecosystem services. Other work by Rayfield et al. (2019, unpublished) has extended the analysis of connectivity to the whole of the Saint Lawrence Lowlands. Beyond connectivity the project of Monteregie Connection, spearheaded by an interdisciplinary team of researchers from McGill University, studied the provision of ecosystem services in the region and produced important knowledge for conservation planning. \\
